# Methodology

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(here)
library(osmdata)
library(sf)
library(tidyverse)
library(ggmap)
library(leaflet)
library(tidycensus)
library(qwraps2)
library(knitrProgressBar)
basetiles <- providers$CartoDB.Positron
this_crs <- 2227 # EPSG:2227 – NAD83 / California zone 3 (ftUS)
bb <- getbb("Alameda County, California", format_out = "polygon")
options(tigris_use_cache = TRUE)
```



We constructed a dataset on which to estimate park activity location choices for
a sample of observed trips in Alameda County, California. Alameda County is one
of the seven counties that constitutes the San Francisco Bay Area metropolitan
region in California. Alameda is the seventh most populous county in California
with a population of 1.5 million [@alamedafacts], and has 14 incorporated cities
and several unincorporated communities. It is an economically and ethnically
diverse county and hence it was an attractive area to use for this study. The
racial makeup of Alameda County was (49.7%) White, (11.2%) African American,
(1.0%) Native American, (38.7%) Asian, (1.0%) Pacific Islander, and (22.4%
Hispanic or Latino (of any race). Alameda County has a diverse set of parks,
ranging from local small community parks, urban and transit-accessible parks
like the Lake Merritt Recreational area, accessible coastal access, and suburban
recreational areas like Lake Chabot.


## Data
We constructed an analysis dataset from a publicly-available parks polygons
layer, a commercially acquired passive device origin-destination table
representing trips between the parks and home block groups, and American
Community Survey data for the home block groups.

We obtained a polygons shapefile layer representing open spaces in Alameda
County, California from the California Protected Areas Database [@cpad2019].
This dataset was selected because it included multiple different types of open
space including local and state parks, traditional green spaces as well as
wildlife refuges and other facilities that can be used for recreation. We
removed facilities that did not allow open access to the public (such as the
Oakland Zoo) and facilities whose boundaries conflated with freeway right-of-way
– this prevents trips through the park from being conflated with park trips in
the passive origin-destination data. 

```{r parks}
not_parks <- c("3455", "15886", "13243")
parks <- st_read(here("data/bayarea_parks.geojson"), quiet = TRUE) %>%
  filter(COUNTY == "Alameda") %>%
  transmute(
    id = as.character(UNIT_ID), name = UNIT_NAME, 
    access = factor(ACCESS_TYP, levels = c("Open Access", "No Public Access", 
                                           "Restricted Access")),
    acres = ACRES, type = DES_TP) %>%
  filter(!id %in% not_parks) %>%
  filter(access == "Open Access") %>%
  select(id, access, acres, type)  %>%
  st_transform(this_crs)
```

From this initial parks polygons dataset, we obtained park attribute information
through OpenStreetMap (OSM) using the `osmdata` package for R [@osmdata]. Three 
attribute elements are considered in this analysis. First, we identify playgrounds
using OSM facilities given a `leisure = playground` tag. The tagged facilities can
be either polygon or point objects; in the former case we use the polygon centroid
to determine the location of the playground.

```{r playgrounds, warning = FALSE}
playground_osm <- opq(bb) %>% # specify boundary for query
  add_osm_feature(key = "leisure", value = "playground") %>% # specify which kinds of data we want
  osmdata_sf() %>% # get a list of sf data frames for these tags
  trim_osmdata(bb, exclude = TRUE)

playgrounds <- rbind(
  # convert polygons to centroids
  playground_osm$osm_polygons %>%
    st_transform(this_crs) %>%
    st_centroid() %>%
    select(osm_id),
  # get points
  playground_osm$osm_points %>%
    st_transform(this_crs) %>%
    select(osm_id)
  # no multipolygons or polylines that need to be included
) %>%
  mutate(playground = TRUE)
```

Second, we consider sport fields of various kinds identified with the OSM
`leisure = pitch` tag. This tag has an additional attribute describing the sport
the field is designed for, which we retain in a consolidated manner. Soccer and
American football fields are considered in a single category, and baseball and
softball fields are combined as well. Basketball, tennis, and volleyball courts
are kept as distinct categories, with all other sport-specific fields combined
into a single "other". Golf courses are discarded. As with playgrounds, polygon 
field and court objects are converted into points at the polygon centroid.

```{r pitches}
pitches_osm <- opq(bb) %>% # specify boundary for query
  add_osm_feature(key = "leisure", value = "pitch") %>% 
  osmdata_sf() %>% 
  trim_osmdata(bb, exclude = TRUE)

pitches <- rbind(
  # convert polygons to centroids
  pitches_osm$osm_polygons %>%
    st_transform(this_crs) %>%
    st_centroid() %>%
    select(osm_id, sport),
  # get points
  pitches_osm$osm_points %>%
    st_transform(this_crs) %>%
    select(osm_id, sport)
  # no multipolygons or polylines that need to be included
) %>%
  filter(sport != "golf") %>%
  mutate(
    sport = case_when(
      grepl("football", sport) | grepl("soccer", sport) ~ "football / soccer",
      grepl("baseball", sport) | grepl("softball", sport) ~ "baseball",
      grepl("basketball", sport) ~ "basketball",
      grepl("tennis", sport) ~ "tennis",
      grepl("volleyball", sport) ~ "volleyball",
      TRUE ~ "other"
    )
  )
```

Finally, we identified trails and footpaths using the `path`, `cycleway`, and
`footway` values of the `highway` tag. A visual inspection of the resulting data
revealed that the large preponderance of sidewalks and cycling trails within parks
in Alameda County are identified properly with these variables. Trails are
represented in OSM as polylines, or as polygons if they form a complete circle. 
In the latter case, we converted the polygon boundary into an explicit polyline object.

```{r trails}
trails_osm <- opq(bb) %>% # specify boundary for query
  add_osm_feature(key = "highway", value = c("footway", "cycleway", "path")) %>% 
  osmdata_sf() %>% 
  trim_osmdata(bb, exclude = FALSE)

trails <- rbind(
  # points are nodes, so we can skip them.
  # lines
  trails_osm$osm_lines %>%
    st_transform(this_crs) %>%
    select(osm_id, bicycle, foot, horse),
  
  # circular paths get converted to polygons, so we need to 
  # turn them back into lines.
  trails_osm$osm_polygons %>%
    st_transform(this_crs) %>%
    st_boundary() %>%
    select(osm_id, bicycle, foot, horse)
) %>%
  mutate(length = st_length(.))
```

It is possible for each of these facilities to exist outside the context of 
a public park. For example, many private apartment complexes have playgrounds
and high schools will have sports facilities that are not necessarily open to
the general public. We spatially matched the OSM amenities data and retained
only those facilities that intersected with the CPAD open spaces database 
identified earlier.  

```{r attributed-parks}
attributed_parks <- parks %>%  
  # compute yeo-johnson transformation on acres
  mutate(yj_acres = VGAM::yeo.johnson(acres, lambda = 0)) %>%
  # determine if a playground point is inside the park polygon boundary
  st_join(playgrounds %>% select(playground), st_intersects)  %>%
  # a park with multiple playgrounds will join multiple times, so we need to 
  # summarize them back down.
  group_by(id) %>% slice(1) %>% ungroup() %>%
  mutate(playground = ifelse(is.na(playground), F, playground)) %>%
  
  # determine if the pitch points are inside the park polygon boundaries
  st_join(pitches %>% select(sport)) %>%
  # multiple pitches in a park result in repeated rows. 
  group_by(id) %>% slice(1) %>% ungroup() %>%
  mutate(pitch = ifelse(is.na(sport), F, T)) %>%
  select(-sport) %>%
  
  # determine if the trails are inside the park polygon boundaries
  st_join(trails %>% select(trail = osm_id), st_intersects) %>%
  # multiple pitches in a park result in repeated rows. 
  group_by(id) %>% slice(1) %>% ungroup() %>%
  mutate(trail = ifelse(is.na(trail), F, T))
```


```{r parks-map, fig.cap="Location of e-scooter users in Provo, Utah.", warning=FALSE, cache = FALSE}
if(knitr::is_latex_output()) {
  alameda_back <- get_googlemap(c(lon = -122.1300334, lat = 37.69427), 
                                zoom = 9, maptype = "terrain", color = "bw") 
  plot(st_transform(parks, 3857)[,"access"], bgMap = alameda_back)
} else {
  leaflet() %>%
    addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
    addPolygons(data = parks %>% st_transform(4326), group = "Parks", color = "green") %>%
    addCircleMarkers(data = playgrounds %>% st_transform(4326), 
                     group = "Playgrounds", color = "blue", radius = 2) %>%
    addCircleMarkers(data = pitches %>% st_transform(4326), 
                     group = "Sport Fields", color = "red", radius = 2) %>%
    addPolylines(data = trails %>% st_transform(4326), 
                 group = "Trails", color = "black") %>%
    addLayersControl(overlayGroups = c("Parks", "Playgrounds", "Sport Fields", "Trails"))
}
```


We provided the park boundaries layer to a commercial firm, StreetLight Data
Inc., which develops and resells origin-destination matrices derived from
passive device location data. The provider employs a proprietary data processing
engine (called Route Science) to algorithmically transform observed device
location data points (the provider uses in-vehicle GPS units and mobile device
LBS) over time into contextualized, normalized, and aggregated travel patterns.
From these travel patterns, the Route Science processing algorithms infer likely
home Census block group locations for composite groups of people and converts
raw location data points into trip origin and destination points [@Pan2006;
@Friedrich2010].

```{r park_flows}
park_flows <- read_csv(here("data/Master_BlockGroups_Final071519.csv"), 
                       col_types = cols()) %>% 
  # need to make block group id properly formatted.
  transmute(
    park = as.character(`Zone Name`),
    home = paste0("0", round(`Block Group ID`)),
    county = substr(home, 0, 5),
    attractions = `Zone Traffic (StL Index)`,
    d_share = `Percent by Home Location`
  ) %>%
  
  # only keep bgs in the area
  filter(county %in% c("06001")) %>%
  # and only keep parks that we are looking at
  filter(park %in% parks$id) %>%
  
  mutate(flow = attractions * d_share) %>%
  group_by(home) %>%
  mutate(
    productions = sum(flow),
    o_share = flow / productions
  ) 
``` 

For each park polygon, the firm returned a population-weighted estimate of how
many devices were observed by home location block group over several months in
the period between May 2018 and October 2018. We transformed this table such
that it represented the weighted unique devices traveling between block groups
and parks. We discarded home location block groups outside of Alameda County;
the imputed home locations can be far away from the study area for a small
amount of trips and are unlikely to represent common or repeated park
activities.

```{r parks-table, results="asis"}
attributed_parks <- attributed_parks %>% group_by(type) %>%
  left_join(park_flows %>% group_by(park) %>% slice(1) %>% 
              select(park, attractions), by = c("id" = "park")) %>%
  mutate(attractions = ifelse(is.na(attractions), 0, attractions)) 
```

Table \ref{tab:parks-table} presents descriptive statistics
on the `r nrow(parks)` parks assembled for this study, grouped according to the
park type as defined on CPAD. A little more than half of the parks have
identified paths, while around 40% of the identified parks have playgrounds and
sport fields.

```{r park_attributes, results="asis"}
park_attributes <- datasummary_balance(
  ~ type,  
  data = attributed_parks %>% 
    select(Access = access, Acres = acres, type, Playground = playground, 
           `Sport Field` = pitch, Trail = trail, 
           `Mobile Devices` =attractions), 
  dinm = FALSE,
  title = "Park Summary Statistics",
  output = "kableExtra")
```



In order to understand the demographic makeup of the home block groups and
potentially the characteristics of the people who make each trip, we obtained
2013-2017 five-year data aggregations from the American Community Survey 
using the `tidycensus` [@Walker2019] interface to the
Census API for several key demographic and built environment variables: the
share of individuals by ethnic group, the share of households by income level,
household median income, and the housing unit density. An important attribute of
the choice model is the distance from the home block group to the park boundary.
Because we have no information on where in the block group a home is actually
located, we use the population-weighted block group centroid published by the
Census Bureau as the location for all homes in the block group. We then measured
the network-based distance between the park and the home block group centroid
using a walk network derived from OpenStreetMap using the `networkx` package
for Python [@networkx],


```{r acs, message = FALSE}
variables <- c(
  "population" = "B25008_001", # TOTAL POPULATION IN OCCUPIED HOUSING UNITS BY TENURE
  "housing_units" = "B25001_001", # HOUSING UNITS
  "households" = "B19001_001", #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  #Estimate!!Total!!Female!!Worked in the past 12 months!!Usually worked 35 or more hours per week
  # RACE
  "black" = "B02001_003",
  "asian" = "B02001_005",
  "pacific" = "B02001_006",
  "nativeam" = "B02001_004",
  # HISPANIC OR LATINO ORIGIN BY SPECIFIC ORIGIN
  # The number of hispanic individuals needs to be drawn from a different table.
  "hispanic" = "B03001_003",
  #MEDIAN HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  "income" = "B19013_001",
  #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  "inc_0010" = "B19001_002",  "inc_1015" = "B19001_003", "inc_1520" = "B19001_004",
  "inc_2025" = "B19001_005", "inc_2530" = "B19001_006", "inc_3035" = "B19001_007",
  "inc_125"  = "B19001_015", "inc_150"  = "B19001_016", "inc_200"  = "B19001_017"
)

acs <- get_acs(geography = "block group", variables = variables, year = 2017,
               state = "CA", county = "001", geometry = TRUE) %>%
  select(-moe) %>%
  spread(variable, estimate) %>%
  mutate(area = as.numeric(st_area(geometry) * 1e-6)) %>%
  # area is in m^2, change to km^2
  transmute(
    geoid = GEOID,
    population, households, housing_units, 
    density = households / area,
    income,
    # many of the variables come in raw counts, but we want to consider
    # them as shares of a relevant denominator.
    lowincome    = 100 * (inc_0010 + inc_1015 + inc_1520 + inc_2530 +
                            inc_3035) / households,
    highincome   = 100 * (inc_125 + inc_150 + inc_200) / households,
    black        = 100 * black / population,
    asian        = 100 * asian / population,
    other        = 100 * (nativeam + pacific) / population,
    white_hisp   = 100 - black - asian - other
  )
```


```{r net_distances}
# distances computed in py/shortest_paths.py
path_files <- dir("data/shortest_paths/", full.names = TRUE)
distance_df <- lapply(path_files, function(file) {
  read_csv(file, col_types = list(geoid = col_character(), park_id = col_character())) %>%
  mutate(
    yj_distance = VGAM::yeo.johnson(distance, lambda = 0),
    yj_euc_dist = VGAM::yeo.johnson(euc_dist, lambda = 0)
  )
}) %>%
  bind_rows() %>%
  rename(home = geoid, park = park_id) %>%
  mutate(home = sprintf("%012s", home))
```



```{r acs-table}
tracts_data_index <- tibble(
  variable =  c("density", 
    "income", "lowincome", "highincome",
    "black", "asian", "other")
) %>%
  mutate(order = row_number())
acs %>%
  st_set_geometry(NULL) %>%
  filter(population > 0) %>%
  tbl_df() %>%
  dplyr::select(tracts_data_index$variable) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(
    `msd` = paste(median_iqr(value, na_rm = TRUE, show_n = "never", digits = 1))
  ) %>%
  inner_join(tracts_data_index, by = "variable") %>%
  arrange(order) %>%
  mutate(
    variable = c(
      "Density: Households per square kilometer",
      "Income: Median tract income",
      "Low Income: Share of households making less than $35k",
      "High Income: Share of households making more than $125k",
      "Black: Share of population who is black",
      "Asian: Share of population who is Asian",
      "Other: Share of population who belong to other minority groups")
  ) %>%
  dplyr::select(-order) %>%
  knitr::kable(col.names = c("", "Median (IQR)"), booktabs = TRUE,
        digits = 1,
        caption = "Descriptive Statistics for Residence Block Groups")
```


## Model

In random utility choice theory, if an individual living in block group $n$
wishes to make a park trip, the probability that the individual will choose
park $i$ from the set of all parks $J$ can be described as a ratio of the
park's measurable utility $V_{ni}$ to the sum of the utilities for all parks
in the set. In the common destination choice framework we apply a
multinomial logit model [@McFadden1974, @Recker1978],
\begin{equation}\label{eq:p}
   P_{ni} = \frac{\exp(V_{ni})}{\sum_{j \in J}\exp(V_{nj})}
\end{equation}
where the measurable utility $V_{ni}$ is a linear-in-parameters function of 
the destination attributes. 
\begin{equation}\label{eq:V}
V_{ni} = X\beta
\end{equation}
where $\beta$ is a vector of estimable coefficients giving the relative utility
(or disutility) of that attribute to the choice maker, all else equal. It is
possible to add amenities of the park or the journey to the utility
equation. However, as the number of alternatives is large, it is impractical to
consider alternative-specific constants or coefficients and therefore not
possible to include attributes of the home block group or traveler $n$ directly.
We can, however, segment the data and estimate different distance and size
parameters for different segments to observe heterogeneity in the utility
parameters between different socioeconomic groups.

The logarithm of the sum in the denominator of Equation \ref{eq:p} (called the
logsum) provides a measure of the consumer surplus of the choice set
[@Williams1977a],
\begin{equation}\label{eq:logsum}
CS_n = \ln{{\sum_{j \in J}\exp(V_{nj})}} + C
\end{equation}
where $C$ is a constant indicating an unknown absolute value. But comparing the
relative logsum values across choice makers, $CS_n - CS_{n-1}$ gives an
indication of which choice maker has a more valuable choice set. Or, in this
case of a park destination choice model, which choice maker has better access to
parks. Such a "utility-based" accessibility term is thus continuously defined,
dervied directly from choice theory, and can contain multiple dimensions of the
attributes of the choice [@Handy1997;@Dong2006].


```{r params}
set.seed(42)
n_obs <- 20000
n_alts <- 10
n_flow <- sum(park_flows$flow)
ll0e <- sum(n_obs *.8 * log(1/n_alts))
```


```{r mydata}
mydata <- park_flows %>%
  ungroup() %>%
  mutate(weight = flow / sum(flow)) %>%
  sample_n(n_obs, replace = TRUE, weight = weight) %>%
  transmute(id = row_number(), home, alt_0 = park, 
            validation = sample(c(TRUE,FALSE), n(), TRUE, prob = c(0.2, 0.8)))
```

In the most typical cases, researchers estimate the utility coefficients for
destination choice models from household travel surveys. As we have no knowledge
of an appropriate survey on park access, we need to synthesize a suitable
estimation data set. We do this by sampling
`r n_obs` random discrete device origin-destination pairs from the commercial
passive data matrix, weighted by the volume of the flows. This corresponds to a
`r round(n_obs / n_flow * 100, 1)`\% sample of all the observed device
origin-destination pairs.

The sampled origin-destination pair gives the home location as well as the
"chosen" alternative for a synthetic person. In principle the individual's
choice set contains all the parks in our dataset; in practice it can be
difficult to estimate choice models with so many alternatives 
($|J| = `r nrow(parks)`$). For this reason we randomly sample `r n_alts` additional parks
to serve as the non-chosen alternatives for our synthetic choice maker. Such
random sampling of alternatives reduces the efficiency of the estimated
coefficients but the coefficients remain unbiased [@train2009]. As the model has
no alternative-specific constants, the standard likelihood comparison statistic
against the market shares model $\rho^2$ is not computable. We instead use the
likelihood comparison against the equal shares model $\rho_0^2$.

```{r sampled_parks}
sampled_parks <- lapply(1:n_obs, function(i){
  sample(attributed_parks$id, n_alts)
}) %>%
  unlist() %>%
  matrix(ncol = n_alts) %>%
  as_tibble(.name_repair = ~ str_c("alt", 1:n_alts, sep = "_"))

logitdata <- mydata %>%
  bind_cols(sampled_parks) %>%
  gather(key = "alt", value = "park", -id, -home, -validation) %>%
  mutate(chosen = alt == "alt_0") %>%
  arrange(id, alt) %>%
  
  # append distances
  inner_join(distance_df, by = c("home", "park")) %>%
  
  # append park attributes
  left_join(attributed_parks, by = c("park" = "id")) %>%
  
  # append block group attributes
  left_join(acs %>% select(geoid, density:other) %>% st_set_geometry(NULL),
            by = c("home" = "geoid")
   
)

write_rds(logitdata, "data/logitdata.rds")
```


The resulting analysis dataset therefore contains `r n_obs` choice makers that
select between `r n_alts + 1` parks including the park they were observed to
choose; the measured distance between the choice maker's block group and all
parks in the choice set; and the acreage of each park in the choice set. We hold
out a random sample of approximately 20% of choice makers for validation
purposes. We use the `mlogit` package for R [@mlogit; @R] to estimate the
multinomial logit models. 

