# Methodology {#methodology}

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)

inline_hook <- function(x) {
  if (is.numeric(x)) {
    format(x, digits = 3, big.mark = ",")
  } else x
}
knitr::knit_hooks$set(inline = inline_hook)

if(knitr::is_latex_output()){
  knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
}

library(here)
library(osmdata)
library(sf)
library(tidyverse)
library(ggmap)
library(leaflet)
library(tidycensus)
library(modelsummary)
library(knitrProgressBar)
basetiles <- providers$CartoDB.Positron
this_crs <- 2227 # EPSG:2227 – NAD83 / California zone 3 (ftUS)
bb <- getbb("Alameda County, California", format_out = "polygon")
options(tigris_use_cache = TRUE)
options(dplyr.summarise.inform = FALSE)

source("R/methods.R")
```



We first constructed a dataset on which to estimate park activity location
choices for a sample of observed trips in Alameda County, California. Then we
estimated a set of models... Finally, we applied the results of those models to
estimated the change in consumer surplus...

## Study area

Alameda County is one of nine counties that constitute the San Francisco Bay
Area metropolitan region in California. Alameda is the seventh most populous
county in California with a population of 1.5 million [@alamedafacts], and has
14 incorporated cities and several unincorporated communities. It is an
economically and ethnically diverse county and hence it was an attractive area
to use for this study. The racial makeup of Alameda County was (49.7%) White,
(11.2%) African American, (1.0%) Native American, (38.7%) Asian, (1.0%) Pacific
Islander, and (22.4%) Hispanic or Latino (of any race). Alameda County has a
diverse set of parks, ranging from local small community parks, urban and
transit-accessible parks like the Lake Merritt Recreational area, accessible
coastal access, and suburban recreational areas like Lake Chabot.

## Data
We estimated a utility model using a dataset constructed from a
publicly-available parks polygons layer, a commercially acquired passive device
origin-destination table representing trips between the parks and home block
groups, and American Community Survey data for the home block groups. We then
applied that model to a dataset representing open street installations that were
implemented in response to the COVID-19 pandemic.

### Model estimation data

We obtained a polygons shapefile layer representing open spaces in Alameda
County, California from the California Protected Areas Database [@cpad2019].
This dataset was selected because it included multiple different types of open
space including local and state parks, traditional green spaces as well as
wildlife refuges and other facilities that can be used for recreation. We
removed facilities that did not allow open access to the public (such as the
Oakland Zoo) and facilities whose boundaries conflated with freeway right-of-way
– this prevents trips through the park from being conflated with park trips in
the passive origin-destination data.

```{r parks, echo=FALSE}
not_parks <- c("3455", "15886", "13243")
parks <- st_read(here("data/bayarea_parks.geojson"), quiet = TRUE) %>%
  filter(COUNTY == "Alameda") %>%
  transmute(
    id = as.character(UNIT_ID), name = UNIT_NAME, 
    access = factor(ACCESS_TYP, levels = c("Open Access", "No Public Access", 
                                           "Restricted Access")),
    acres = ACRES, type = DES_TP) %>%
  filter(!id %in% not_parks) %>%
  filter(access == "Open Access") %>%
  select(id, access, acres, type)  %>%
  st_transform(this_crs)
```

From this initial parks polygons dataset, we obtained park attribute information
through OpenStreetMap (OSM) using the `osmdata` package for R [@osmdata]. Three
attribute elements are considered in this analysis. First, we identify
playgrounds using OSM facilities given a `leisure = playground` tag. The tagged
facilities can be either polygon or point objects; in the former case we use the
polygon centroid to determine the location of the playground.

```{r playgrounds, warning = FALSE, echo= FALSE}
playground_osm <- opq(bb) %>% # specify boundary for query
  add_osm_feature(key = "leisure", value = "playground") %>% # specify which kinds of data we want
  osmdata_sf() %>% # get a list of sf data frames for these tags
  trim_osmdata(bb, exclude = TRUE)

playgrounds <- rbind(
  # convert polygons to centroids
  playground_osm$osm_polygons %>%
    st_transform(this_crs) %>%
    st_centroid() %>%
    select(osm_id),
  # get points
  playground_osm$osm_points %>%
    st_transform(this_crs) %>%
    select(osm_id)
  # no multipolygons or polylines that need to be included
) %>%
  mutate(playground = TRUE)
```

Second, we consider sport fields of various kinds identified with the OSM
`leisure = pitch` tag. This tag has an additional attribute describing the sport
the field is designed for, which we retain in a consolidated manner. Soccer and
American football fields are considered in a single category, and baseball and
softball fields are combined as well. Basketball, tennis, and volleyball courts
are kept as distinct categories, with all other sport-specific fields combined
into a single "other." Golf courses are discarded. As with playgrounds, polygon
field and court objects are converted into points at the polygon centroid.

```{r pitches, warning = FALSE, echo=FALSE}
pitches_osm <- opq(bb) %>% # specify boundary for query
  add_osm_feature(key = "leisure", value = "pitch") %>% 
  osmdata_sf() %>% 
  trim_osmdata(bb, exclude = TRUE)

pitches <- rbind(
  # convert polygons to centroids
  pitches_osm$osm_polygons %>%
    st_transform(this_crs) %>%
    st_centroid() %>%
    select(osm_id, sport),
  # get points
  pitches_osm$osm_points %>%
    st_transform(this_crs) %>%
    select(osm_id, sport)
  # no multipolygons or polylines that need to be included
) %>%
  filter(sport != "golf") %>%
  mutate(
    sport = case_when(
      grepl("football", sport) | grepl("soccer", sport) ~ "football / soccer",
      grepl("baseball", sport) | grepl("softball", sport) ~ "baseball",
      grepl("basketball", sport) ~ "basketball",
      grepl("tennis", sport) ~ "tennis",
      grepl("volleyball", sport) ~ "volleyball",
      TRUE ~ "other_pitch"
    )
  )
```

Finally, we identified trails and footpaths using the `path`, `cycleway`, and
`footway` values of the `highway` tag. A visual inspection of the resulting data
revealed that the large preponderance of sidewalks and cycling trails within
parks in Alameda County are identified properly with these variables. Trails are
represented in OSM as polylines, or as polygons if they form a complete circle.
In the latter case, we converted the polygon boundary into an explicit polyline
object.

```{r trails, echo = FALSE}
trails_osm <- opq(bb) %>% # specify boundary for query
  add_osm_feature(key = "highway", value = c("footway", "cycleway", "path")) %>% 
  osmdata_sf() %>% 
  trim_osmdata(bb, exclude = FALSE)

trails <- rbind(
  # points are nodes, so we can skip them.
  # lines
  trails_osm$osm_lines %>%
    st_transform(this_crs) %>%
    select(osm_id, bicycle, foot, horse),
  
  # circular paths get converted to polygons, so we need to 
  # turn them back into lines.
  trails_osm$osm_polygons %>%
    st_transform(this_crs) %>%
    st_boundary() %>%
    select(osm_id, bicycle, foot, horse)
) %>%
  mutate(length = st_length(.))
```

It is possible for each of these facilities to exist outside the context of  a
public park. For example, many private apartment complexes have playgrounds and
high schools will have sports facilities that are not necessarily open to the
general public. We spatially matched the OSM amenities data and retained only
those facilities that intersected with the CPAD open spaces database identified
earlier.

```{r attributed-parks, echo=FALSE}
# Append attributes to park frame ===============
parks_with_playgrounds <-  parks %>%  
  # compute yeo-johnson transformation on acres
  # determine if a playground point is inside the park polygon boundary
  st_join(playgrounds %>% select(playground), st_intersects)  %>%
  # a park with multiple playgrounds will join multiple times, so we need to 
  # summarize them back down.
  group_by(id) %>% slice(1) %>% ungroup() %>%
  mutate(playground = ifelse(is.na(playground), F, playground)) %>%
  select(id, playground) %>%
  st_set_geometry(NULL)
  
# determine if the pitch points are inside the park polygon boundaries
parks_with_pitches <- parks %>%
  select(id) %>%
  st_join(pitches %>% select(sport)) %>%
  st_set_geometry(NULL) %>%
  # multiple pitches in a park result in repeated rows. 
  group_by(id, sport) %>% summarise(count = n()) %>%
  pivot_wider(id_cols = id, values_from = count, names_from = sport, 
              values_fill = FALSE) %>%
  mutate(
    pitch = ifelse(`NA` == 1, FALSE, TRUE),
    across(baseball:tennis, ~ifelse(. > 0, TRUE, FALSE)),
  ) %>%
    select(-`NA`)

# determine which walking paths are inside park polgon boundaries
parks_with_trails <- parks %>% 
  # determine if the trails are inside the park polygon boundaries
  st_join(trails %>% select(trail = osm_id), st_intersects) %>%
  # multiple pitches in a park result in repeated rows. 
  group_by(id) %>% slice(1) %>% ungroup() %>%
  mutate(trail = ifelse(is.na(trail), F, T)) %>%
  select(id, trail) %>%
  st_set_geometry(NULL)

# join together and write out =================
attributed_parks <- parks %>%
  # compute yeo-johnson transformation on acres
  mutate(yj_acres = VGAM::yeo.johnson(acres, lambda = 0)) %>%
  left_join(parks_with_playgrounds, by = "id") %>%
  left_join(parks_with_pitches,     by = "id") %>%
  left_join(parks_with_trails,      by = "id") 
```


```{r parks-map, fig.cap="Location of parks in Alameda County.", warning=FALSE, cache = FALSE, echo=FALSE}
dj1 <- wesanderson::wes_palette("Darjeeling1")
if(knitr::is_latex_output()) {
  pbox <- st_bbox(st_transform(parks, 4326))
  alameda_back <- get_map(c(lon = mean(pbox[c(1, 3)]), 
                            lat = mean(pbox[c(2, 4)]) ), 
                          zoom = 10, source = "stamen", color = "bw")
  alameda_back <- ggmap_bbox(alameda_back)
  
  ggmap(alameda_back, extent = "device") + 
    coord_sf(crs = st_crs(3857), 
             xlim = c(-122.29674 , -122.16358), 
             ylim = c( 37.74098, 37.87028), expand = FALSE) + 
    geom_sf(data = parks %>% st_transform(3857), inherit.aes = FALSE,
            fill = dj1[2], lwd = 0)  + 
    theme(axis.line = element_line(color = NA))  + 
    xlab("") + ylab("")
} else {
  leaflet() %>%
    addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
    addPolygons(data = parks %>% st_transform(4326), group = "Parks", color = dj1[2]) %>%
    addCircleMarkers(data = playgrounds %>% st_transform(4326), 
                     group = "Playgrounds", color = dj1[3], radius = 2) %>%
    addCircleMarkers(data = pitches %>% st_transform(4326), 
                     group = "Sport Fields", color = dj1[1], radius = 2) %>%
    addPolylines(data = trails %>% st_transform(4326), 
                 group = "Trails", color = dj1[5]) %>%
    addLayersControl(overlayGroups = c("Parks", "Playgrounds", "Sport Fields", "Trails"))
}
```

We provided the park boundaries layer to a commercial firm, StreetLight Data
Inc., which develops and resells origin-destination matrices derived from
passive device location data. The provider employs a proprietary data processing
engine (called Route Science) to algorithmically transform observed device
location data points (the provider uses in-vehicle GPS units and mobile device
LBS) over time into contextualized, normalized, and aggregated travel patterns.
From these travel patterns, the Route Science processing algorithms infer likely
home Census block group locations for composite groups of people and converts
raw location data points into trip origin and destination points [@Pan2006;
@Friedrich2010].

```{r park_flows, echo=FALSE}
# park flows computed from Streetlight data in R/park_flows.R
park_flows <- read_rds("data/park_flows.rds")
``` 

For each park polygon, the firm returned a population-weighted estimate of how
many devices were observed by home location block group over several months in
the period between May 2018 and October 2018. We transformed this table such
that it represented the weighted unique devices traveling between block groups
and parks. We discarded home location block groups outside of Alameda County;
the imputed home locations can be far away from the study area for a small
amount of trips and are unlikely to represent common or repeated park
activities. Table \@ref(tab:park-attributes) presents descriptive statistics on
the `r nrow(parks)` parks assembled for this study, grouped according to the
park type as defined on CPAD. A little more than half of the parks have
identified paths, while around 40% of the identified parks have playgrounds and
sport fields.

```{r parkstable, echo=FALSE}
attributed_parks <- attributed_parks %>% group_by(type) %>%
  left_join(park_flows %>% group_by(park) %>% slice(1) %>% 
              select(park, attractions), by = c("id" = "park")) %>%
  mutate(attractions = ifelse(is.na(attractions), 0, attractions)) 
write_rds(attributed_parks, "data/attributed_parks.rds")
```


```{r park-attributes, results="asis", echo=FALSE}
datasummary_balance(
  ~ Type,  
  data = attributed_parks %>% 
    transmute(Access = access, Acres = acres, 
           Type = ifelse(grepl("Recreation Area", type), "Recreation Area" ,type),
           Playground = playground, 
           `Any Sport Field` = pitch, 
           `Football / Soccer` = `football / soccer`,
           `Baseball` = baseball, `Basketball` = basketball,
           `Tennis` = tennis, `Volleyball` = volleyball,
           Trail = trail, 
           `Mobile Devices` =attractions), 
  dinm = FALSE, caption = "Park Summary Statistics", booktabs = TRUE)
```

In order to understand the demographic makeup of the home block groups and
potentially the characteristics of the people who make each trip, we obtained
2013-2017 five-year data aggregations from the American Community Survey  using
the `tidycensus` [@Walker2019] interface to the Census API for several key
demographic and built environment variables: the share of individuals by race,
the share of households by income level, household median income, the share of
households with children under 6 years old, and the household density. An
important attribute of the choice model is the distance from the home block
group to the park boundary. Because we have no information on where in the block
group a home is actually located, we use the population-weighted block group
centroid published by the Census Bureau as the location for all homes in the
block group. We then measured the network-based distance between the park and
the home block group centroid using a walk network derived from OpenStreetMap
using the `networkx` package for Python [@networkx],

```{r acs, message = FALSE, echo=FALSE}
variables <- c(
  "population" = "B02001_001", # TOTAL: RACE
  "housing_units" = "B25001_001", # HOUSING UNITS
  "households" = "B19001_001", #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  # Hispanic or Latino Origin by Race
  "white" = "B03002_003",
  "black" = "B03002_004",
  "asian" = "B03002_006",
  "hispanic" = "B03002_012",
  #MEDIAN HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  "income" = "B19013_001",
  # FAMILY TYPE BY PRESENCE AND AGE OF RELATED CHILDREN
  "children_c06" = "B11004_004", # married under 6 only
  "children_c6+" = "B11004_005", # married under 6 and older
  "children_m06" = "B11004_011", # male under 6 only
  "children_m6+" = "B11004_012", # male under 6 and older
  "children_f06" = "B11004_017", # female under 6 only
  "children_f6+" = "B11004_018", # female under 6 and older
  #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  "inc_0010" = "B19001_002",  "inc_1015" = "B19001_003", "inc_1520" = "B19001_004",
  "inc_2025" = "B19001_005", "inc_2530" = "B19001_006", "inc_3035" = "B19001_007",
  "inc_125"  = "B19001_015", "inc_150"  = "B19001_016", "inc_200"  = "B19001_017"
)

acs <- get_acs(geography = "block group", variables = variables,
               state = "CA", county = "Alameda", geometry = TRUE) %>%
  select(-moe) %>%
  spread(variable, estimate) %>%
  # area is in m^2, change to km^2
  mutate(area = as.numeric(st_area(geometry) * 1e-6)) %>%
  transmute(
    geoid = GEOID,
    group = 1,
    population, households, housing_units, 
    density = households / area,
    income,
    # many of the variables come in raw counts, but we want to consider
    # them as shares of a relevant denominator.
    children = 100 * ( children_c06 + `children_c6+` + 
                         children_m06 + `children_m6+` + 
                         children_f06 + `children_f6+`) / households,
    lowincome    = 100 * (inc_0010 + inc_1015 + inc_1520 + inc_2530 +
                            inc_3035) / households,
    highincome   = 100 * (inc_125 + inc_150 + inc_200) / households,
    black        = 100 * black / population,
    asian        = 100 * asian / population,
    hispanic     = 100 * hispanic / population,
    white        = 100 * white / population
  ) %>%
  filter(population > 0)

write_rds(acs, "data/blockgroups.rds")
```


```{r net_distances, echo=FALSE}
# distances computed in py/shortest_paths.py
shortest_path_dir <- "data/shortest_paths"
if(!dir.exists(shortest_path_dir)) {
  # CTV: This bit doesn't work on my computer (can't extract 7z
  # files from code? So I just extracted them from the 7zip utility
  # within Windows Explorer prior to running)
  system2("7z", args = c("x", "data/shortest_paths.7z", "-odata/"))
}
path_files <- dir(shortest_path_dir, full.names = TRUE)
distance_df <- lapply(path_files, function(file) {
  read_csv(file, col_types = list(geoid = col_character(), park_id = col_character())) %>%
  mutate(
    yj_distance = VGAM::yeo.johnson(distance, lambda = 0),
    yj_euc_dist = VGAM::yeo.johnson(euc_dist, lambda = 0)
  )
}) %>%
  bind_rows() %>%
  rename(home = geoid, park = park_id) %>%
  mutate(home = str_pad(home, width = 12, side = "left", pad = "0"))
write_rds(distance_df, "data/distances.rds")
```



```{r acs-table, echo=FALSE}
acs_for_table <- acs %>% select(
    "Density: Households per square kilometer" = density,
    "Income: Median tract income" = income,
    "Low Income: Share of households making less than $35k" = lowincome,
    "High Income: Share of households making more than $125k" = highincome,
    "Children: Share of households with children under 6" = children,
    "Black: Share of population who is Black" = black,
    "Asian: Share of population who is Asian" = asian,
    "Hispanic: Share of population who is Hispanic*" = hispanic,
    "White: Share of population who is White" = white)

datasummary_skim(acs_for_table, title = "Block Group Summary Statistics",
  histogram = !knitr::is_latex_output()) %>%
  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
  kableExtra::column_spec(1, width = "4cm") %>%
  kableExtra::footnote(symbol = "Hispanic indicates Hispanic individuals of all races; non-Hispanic individuals report a single race alone.") 
```

### Model application data

```{r closed-streets, echo=FALSE}
slowstreets <- st_read("data/slow_streets.geojson", quiet = TRUE)  %>%
  st_transform(this_crs) 

street_parks <- slowstreets %>%
  st_buffer(dist = 25)  %>% # 50 foot wide park ( 4 lanes of traffic)
  mutate(acres = as.numeric(st_area(.)) / 43560) %>% 
  mutate() %>%
  transmute(
    id = str_c("st", ID), access = "Open Access", acres, type = "Street",
    yj_acres = VGAM::yeo.johnson(acres, lambda = 0),
    playground = FALSE, baseball = FALSE, basketball = FALSE, 
    `football / soccer` = FALSE, other_pitch = FALSE, volleyball = FALSE,
    tennis = FALSE, pitch = FALSE, trail = TRUE, attractions = NA
  )
  
write_rds(street_parks, "data/street_parks.rds")
```

We compiled a list of streets in Berkeley, Oakland, and Alameda that were
converted to public open space from each city's respective websites
[@city_of_alameda_slow_2020; @city_of_oakland_oakland_2020;
@city_of_berkeley_berkeley_2020] and referred to the "Shifting Streets" COVID-19
mobility dataset [@slowstreets] to determine whether other cities and places
within Alameda County has similar Open Streets projects (as best as we could
determine, they did not). Based on the information we gathered from these
sources, `r nrow(street_parks)` individual streets were converted; these streets
represent `r as.numeric(sum(st_length(slowstreets))) / 5280` total miles across
the cities of Berkeley, Oakland, and Alameda. For the purposes of this analysis,
we represent each opened street as a single "park" without any sport facilities
or playgrounds, but with a trail / walking path. The database provides the
opened streets as polyline objects; we assert a 25-foot buffer around the line
to represent a polygon with a measurable area. The 25-foot buffer effectively
counts one vehicle lane and one shoulder parking lane in each direction as
converted to "park" space. Finally, we measure the network-based distance from
each population-weighted block group centroid to the nearest boundary of each
new open space facility created by this policy.

## Model estimation

In random utility choice theory, if an individual living in block group $n$
wishes to make a park trip, the probability that the individual will choose park
$i$ from the set of all parks $J$ can be described as a ratio of the park's
measurable utility $V_{ni}$ to the sum of the utilities for all parks in the
set. In the common destination choice framework we apply a multinomial logit
model [@McFadden1974, @Recker1978],
\begin{equation}\label{eq:p}
   P_{ni} = \frac{\exp(V_{ni})}{\sum_{j \in J}\exp(V_{nj})}
\end{equation}
where the measurable utility $V_{ni}$ is a linear-in-parameters function of 
the destination attributes. 
\begin{equation}\label{eq:V}
V_{ni} = X\beta
\end{equation}
where $\beta$ is a vector of estimable coefficients giving the relative utility
(or disutility) of that attribute to the choice maker, all else equal. It is
possible to add amenities of the park or the journey to the utility equation.
However, as the number of alternatives is large, it is impractical to consider
alternative-specific constants or coefficients and therefore not possible to
include attributes of the home block group or traveler $n$ directly. We can,
however, segment the data and estimate different distance and size parameters
for different segments to observe heterogeneity in the utility parameters
between different socioeconomic groups.

The logarithm of the sum in the denominator of Equation \ref{eq:p} (called the
logsum) provides a measure of the consumer surplus of the choice set enjoyed by
person $n$ [@Williams1977a],
\begin{equation}
CS_n = \ln{{\sum_{j \in J}\exp(V_{nj})}} + C
  (\#eq:logsum)
\end{equation}
where $C$ is a constant indicating an unknown absolute value;  the difference of
logsum values in two different scenarios eliminates $C$. Additionally, dividing
the difference in logsum from choice set $J$ and choice set $J'$ by a cost
coefficient $\beta$
\begin{equation}
\delta CS_n = (\ln{\sum_{j \in J'}\exp(V_{nj})} - \ln{\sum_{j \in J}\exp(V_{nj})})/\beta
  (\#eq:deltalogsum)
\end{equation}
gives an estimate of the benefit received by person $n$ in monetary terms. Thus,
such a "utility-based" accessibility term is continuously defined, contains
multiple dimensions of the attributes of the choice, and can be evaluated in
monetary terms [@Handy1997;@Dong2006].


```{r params, echo=FALSE}
set.seed(42)
n_obs <- 20000
n_alts <- 10
n_flow <- sum(park_flows$flow)
ll0e <- sum(n_obs *.8 * log(1/n_alts))
```


```{r mydata, echo=FALSE}
mydata <- park_flows %>%
  ungroup() %>%
  mutate(weight = flow / sum(flow)) %>%
  sample_n(n_obs, replace = TRUE, weight = weight) %>%
  transmute(id = row_number(), home, alt_0 = park, 
            validation = sample(c(TRUE,FALSE), n(), TRUE, prob = c(0.2, 0.8)))
```

In the most typical cases, researchers estimate the utility coefficients for
destination choice models from houzsehold travel surveys. For example, the
California add-on to the 2017 National Household Travel Survey could be used for
this purpose for frequent trips like commutes to work and school. However, as a
24-hour trip diary, it is less useful for recreational trips that may take place
less frequently. For better data on park access, we need to synthesize a
suitable estimation data set. We do this by sampling `r n_obs` random discrete
device origin-destination pairs from the commercial passive data matrix,
weighted by the volume of the flows. This corresponds to a `r round(n_obs /
n_flow * 100, 1)`\% sample of all the observed device origin-destination pairs.

The sampled origin-destination pair gives the home location as well as the
"chosen" alternative for a synthetic person. In principle the individual's
choice set contains all the parks in our dataset; in practice it can be
difficult to estimate choice models with so many alternatives  ($|J| = `r
nrow(parks)`$). For this reason we randomly sample `r n_alts` additional parks
to serve as the non-chosen alternatives for our synthetic choice maker. Such
random sampling of alternatives reduces the efficiency of the estimated
coefficients but the coefficients remain unbiased [@train2009]. As the model has
no alternative-specific constants, the standard likelihood comparison statistic
against the market shares model $\rho^2$ is not computable. We instead use the
likelihood comparison against the equal shares model $\rho_0^2$.

```{r sampled_parks, echo=FALSE}
sampled_parks <- lapply(1:n_obs, function(i){
  sample(attributed_parks$id, n_alts)
}) %>%
  unlist() %>%
  matrix(ncol = n_alts) %>%
  as_tibble(.name_repair = ~ str_c("alt", 1:n_alts, sep = "_"))

logitdata <- mydata %>%
  bind_cols(sampled_parks) %>%
  gather(key = "alt", value = "park", -id, -home, -validation) %>%
  mutate(chosen = alt == "alt_0") %>%
  arrange(id, alt) %>%
  
  # append distances
  inner_join(distance_df, by = c("home", "park")) %>%
  
  # append park attributes
  left_join(attributed_parks, by = c("park" = "id")) %>%
  
  # append block group attributes
  left_join(acs %>% select(geoid, density:white) %>% st_set_geometry(NULL),
            by = c("home" = "geoid")
   
)

write_rds(logitdata, "data/logitdata.rds")
```


The resulting analysis dataset therefore contains `r n_obs` choice makers that
select between `r n_alts + 1` parks including the park they were observed to
choose; the measured distance between the choice maker's block group and all
parks in the choice set; and the acreage of each park in the choice set.  We use
the `mlogit` package for R [@mlogit; @R] to estimate the multinomial logit
models.

We estimate a series of models with different utility function definitions, each
estimated on the complete set of synthetic choice makers. The first model  ---
named "Network Distance" --- only considers the distance to the park and the
size of the park in the utility equation. The second and third models include a
vector of park attributes. In the model labeled "Park Attributes," the presence
of any sport field is considered with a single dummy value, and in the "Sport
Detail" model this variable is disaggregated into facilities for different
sports. As described in the [Results](#results) section, the "Sport Detail"
model proved to offer the best model fit.

## Model application

Using the full set of Alameda County parks --- including those added by street
conversion --- we applied the "Sport Detail" non-segmented model to calculate
park choice utilities and utility-based accessibility values for each block
group in Alameda County. As shown with Equation \@ref(eq:deltalogsum), the
difference in utility-based accessibility values with and without the opened
streets is the consumer surplus provided by the policy, converted into a
monetary value by a cost-utility coefficient. The dataset used for this research
does not have any information on travel costs or entrance fees, and such data
would likely not be relevant in the context of urban parks. As a result, there
is no direct link between the utility and a monetary cost in our estimated
models.

As a substitution, we use an estimate of the cost coefficient obtained from the
open-source activity-based travel demand model ActivitySim, which is itself
based on the regional travel model employed by the Metropolitan Transportation
Commission (MTC), the San Francisco Bay regional MPO. ActivitySim uses a cost
coefficient of $-0.6$ divided by the each simulated agent's value of time to
determine destination choices for non-work trips.^[To be precise, this is the
cost coefficient on the mode choice model for social, recreational, and other
trip purposes, which influences destination choice through a logsum-based
impedance term.] In ActivitySim, as in most activity-based travel models, the
value of time is considered to vary with an individual's income, but in this
aggregate destination choice model, an aggregate value of time will suffice. The
average value of time in the synthetic population for the Bay Area is $7.75 per
hour, resulting in a cost coefficient on the destination choice utility of
$-0.215$. Dividing the difference in accessibility logsums by the negative of
this value gives an initial estimate of the monetary value of the policy to each
park user.
