# Results {#results}

```{r estimation-setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(mlogit)
library(modelsummary)
library(broom)
library(kableExtra)
library(sf)
library(ggspatial)
library(leaflet)

inline_hook <- function(x) {
  if (is.numeric(x)) {
    format(x, digits = 3, big.mark = ",")
  } else x
}
knitr::knit_hooks$set(inline = inline_hook)

this_map = c(
    "yj_distance" = "log(Distance)",
    "yj_acres" = "log(Acres)",
    "playgroundTRUE" = "Playground",
    "trailTRUE" = "Trail",
    "pitchTRUE" = "Sport Field",
    "basketballTRUE" = "Basketball",
    "baseballTRUE" = "Baseball",
    "`football / soccer`TRUE" = "Football / Soccer",
    "tennisTRUE" = "Tennis",
    "volleyballTRUE" = "Volleyball",
    "other_pitchTRUE" = "Other Sport"
  )
f <- function(x) format(round(x, 1), big.mark=",")
gm <-  list(
  list("raw" = "nobs",  "clean" = "Num.Obs.", "fmt" = f),
  list("raw" = "AIC",   "clean" = "AIC",      "fmt" = f),
  list("raw" = "logLik","clean" = "Log Likelihood", "fmt" = f),
  list("raw" = "rho20", "clean" = "$\\rho^2_0$", "fmt" = 3)
)
this_crs <- 2227 # EPSG:2227 – NAD83 / California zone 3 (ftUS)
```


We estimated multinomial logit park activity location choice models on the
datset described in the previous section. We applied a Yeo-Johnson
transformation [@Yeo2000] to both the walk distance (in meters) between the
park and the block group centroid, and to the park acreage. The Yeo-Johnson
transformation replicates the constant marginal elasticity of a logarithmic
transformation while avoiding undefined values (e.g., $YJ(0) = 0$). For
simplicity, we call this transformation `log()` in the model results tables.
Using a constant marginal elasticity is better reflective of how people perceive
distances and sizes; a one-mile increase to a trip distance matters more to a
two-mile trip than a ten-mile trip.

```{r base_models}
tar_load(base_models)

detail_lrt <- lrtest(base_models[[2]], base_models[[3]])$`Pr(>Chisq)`[2]
d_ratio_base <- -1 * coefficients(base_models$`Network Distance`)[1] / 
  coefficients(base_models$`Network Distance`)[2] 
d_ratio_sport <- -1 * coefficients(base_models$`Sport Detail`)[1] / 
  coefficients(base_models$`Sport Detail`)[2] 
```

Table \@ref(tab:base-modelsummary) presents the model estimation results for
each estimated model. The "Network Distance" model, which only considers the
distance to the park and the size of the park. results in significant estimated
coefficients of the expected sign. That is, individuals will travel further
distances to reach larger parks. The ratio of the estimated coefficients implies
that on average, people will travel `r d_ratio_base` times further to reach a
park twice as large.

Table \@ref(tab:base-modelsummary) also shows the results of the "Park
Attributes" model, which represents the presence of any sport field with a
single dummy variable, and the "Sport Detail" model, which disaggregates this
variable into facilities for different sports. The value of the size and
distance coefficients change modestly from the "Network Distance" model, with
the implied size to distance trade-off rising to `r d_ratio_sport`. Examining
the two amenities models --- independently and in comparison with each other ---
reveals a few surprising findings. First, it appears that playgrounds and sport
fields in general contribute *negatively* to the choice utility equation. This
is both unintuitive and contradictory to previous findings in this space [e.g.,
@Kinnell2006]. Considering different sports separately, there is a wide variety
of observed response with tennis and volleyball facilities attracting more
trips, and football and basketball facilities attracting fewer, all else equal.
Trails and walking paths give substantive positive utility in both models. The
difference in likelihood statistics between the three models is significant
(likelihood ratio test between Sport Detail and Park Attributes model has
$p$-value  `r detail_lrt`), and so in spite of the curious aggregate findings,
we move forward with this utility specification.

```{r base-modelsummary}
modelsummary(
  base_models, stars = TRUE, output = "kableExtra",
  coef_map = this_map, gof_map = gm,
  estimate = "{estimate} {stars}", 
  statistic = "({std.error})",
  title = "Estimated Model Coefficients",
  note = "Standard errors in parentheses."
) %>%
  scroll_box(width = "100%", box_css = "border: 0px;") 
```



It is worth investigating the heterogeneity in preferences that exist among
populations. Though the income and ethnicity of the synthetic park visitors is
not known, we can segment the estimation dataset based on the socioeconomic
makeup of the visitors' residence block group. The models presented in  Table
\@ref(tab:grouped-modelsummary) were estimated on segments developed in this
manner. Models under the "Race/Ethnicity" heading include a race- and
ethnicity-based segmentation: simulated individuals living in block groups with
more than thirty percent Black residents are included in the ">30% Black" model,
an analogous segmentation for block groups with high Asian and Hispanic
populations are in the ">30% Asian" and ">30% Hispanic" models respectively, and
the "Other" model contains all other block groups. Another set of model
segmentation relies on the share of the population in each block group with
household incomes above or below certain thresholds, and a third relies on the
share of households with children under 6 years old. Again, we use the threshold
definitions largely informed by the distributions in Table \@ref(tab:acs-table).

```{r grouped-modelsummary}
tar_load(grouped_models)
grouped_models_clean <- list(
  "> 30% Asian" = grouped_models$asian,
  "> 30% Black" = grouped_models$black,
  "> 30% Hispanic" = grouped_models$hispanic,
  "Other Eth." = grouped_models$othermin,
  "> 30% Low income" = grouped_models$lowincome,
  "> 50% High income" = grouped_models$highincome,
  "Other Inc." = grouped_models$otherinc,
  "> 25% Children" = grouped_models$children,
  "< 5% Children" = grouped_models$`few children`,
  "Other Children" = grouped_models$otherch
)

modelsummary(
  grouped_models_clean, stars = TRUE, output = "kableExtra",
  coef_map = this_map, gof_map = gm,
  estimate = "{estimate} {stars}", 
  statistic = "({std.error})",
  title = "Estimated Model Coefficients with Block Group Segmentations",
  note = "Standard errors in parentheses."
) %>%
  add_header_above(c(" " = 1, "Race/Ethnicity" = 4, "Income" = 3, "Children" = 3)) %>%
  scroll_box(width = "100%", box_css = "border: 0px;") %>%
  landscape() %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  add_footnote("Simulated individuals segmented based on the share of households meeting the segmentation threshold in the residence block group.")

```

The model estimates in Table \@ref(tab:grouped-modelsummary) reveal noticeable
heterogeneity in the park location choices among visitors from different block
group segments. Park visitors living in block groups with a high proportion of
Black and low-income residents show less affinity for trails and other walkways,
but appear  considerably more sensitive to the distance to a park. Park visitors
living in high-income neighborhoods are less sensitive to the distance to a
park, but receive more utility from certain amenities, in particular trails and
tennis courts. Block groups with a high proportion of Hispanic residents and
residents with children under 6 show the least negative response to playgrounds
of all the segments.
  
Seeing that there is a difference in the response in the model segmentation, it
is also worth considering the role of our segmentation thresholds in these
findings. Figure \@ref(fig:split-plots) shows the estimated coefficients and
confidence intervals for these different amenities at different threshold levels
of segmentation. The threshold level means that at least that percent of the
block group's population falls in that category. The confidence intervals widen
as more observations are excluded from the model. The estimated coefficients for
the different segmentations are identical when the share equals zero, and simply
represent the "Sport Detail" model from Table \@ref(tab:base-modelsummary).

Overall, increasing the segmentation threshold level reveals additional
information about user preferences. First, it should be noted that there is some
inconsistency: for instance, block groups with at least 30% of low income
households show a lower importance of distance than block groups with either 20%
or 40% low income households, though all three estimates are within the same 
confidence intervals. The increasing width of the confidence interval, however,
means it is sometimes difficult to make robust statements. Residents of
block groups with a higher share of Asian individuals or high income households both show
relatively more affinity for tennis courts and trails relative to other groups.
Residents of block groups with increasing shares of Hispanic individuals show
the highest affinity for playgrounds, and park goers from neighborhoods with a
greater share of Black individuals are most sensitive to distance and least
sensitive to park size.

```{r split-plots, fig.cap="Estimated utility coefficients and 95% confidence intervals for park amenities at different socioeconomic threshold levels.", fig.height=7.5, cache = FALSE}
tar_load(split_dat)
tex_labels = c("Asian", "Black", "Children under 6", "Hispanic",  
               "Income $<$ \\$ 35k", "Income $>$ \\$ 125k" )

# make the plot
these_colors <- c(wesanderson::wes_palettes$Darjeeling1, "#046C9A")
split_plots <- ggplot(split_dat, aes(x = share, y = estimate, 
                               ymin = conf.low, ymax = conf.high, 
                color = Segmentation, fill = Segmentation)) + 
  theme_minimal() +
  geom_line() + 
  geom_ribbon(alpha = 0.05, size = 0.1, lty = "dotted") + 
  facet_wrap(~ fct_rev(term), scales = "free_y", ncol = 3)  + 
  xlab("Percent of Block Group Population") + 
  ylab("Estimated coefficient and 95\\% confidence interval")

if(knitr::is_latex_output()) {
  split_plots <- split_plots + 
    scale_color_manual(values = these_colors,labels = tex_labels) +
    scale_fill_manual(values = these_colors, labels = tex_labels) + 
    theme(text = element_text(size=8))
    
  lemon::reposition_legend(split_plots, position = 'top left', panel = "panel-3-4")
} else {
  plotly::ggplotly(split_plots +
                     scale_color_manual(values = these_colors) + 
                     scale_fill_manual(values = these_colors) )
}
```

## Equity Analysis of COVID-19 Street Openings

In this section, we apply the models estimated above to evaluate the benefits of
the street conversion policy in terms of aggregate value determined by the
change in accessibility logsum, as well as the equity of the policy with respect
to different income and ethnic groups. In this analysis, we apply the "Sport Detail"
non-segmented model from Table \@ref(tab:base-modelsummary), as it had the best
fit of these models. 

Figure \@ref(fig:logsumsmap) presents this monetary valuation spatially.
Unsurprisingly, the benefits are concentrated in the block groups surrounding
the opened streets. Most residents of central Oakland see a benefit of somewhere
around \$1, while some zones see an equivalent benefit of as much as \$30. One
property of logsum-based accessibility terms is that there is some benefit given
for simply having more options, whether or not those options are attractive in
any way. In this application, these benefits are small, on the order of 10 cents
for most block groups away from where the street openings occurred.

```{r logsumsmap, fig.cap="Monetary value of street opening to residents based on utility change. Streets converted to pedestrian plazas are shown in black.", cache=FALSE}
bins <- c(0, 0.1, 0.25, 0.5, 1, 2, 10, 35)
mypal <- as.character(wesanderson::wes_palette("Zissou1", n = 7, type = "continuous"))
tar_load(logsums)
tar_load(street_parks)

if(knitr::is_latex_output()) {
  ggplot() +
    annotation_map_tile("cartolight", zoom = 11) +
    coord_sf(crs = st_crs(3857), 
             xlim = c(-122.29674 , -122.16358), 
             ylim = c( 37.74098, 37.87028), expand = FALSE) + 
    geom_sf(data = logsums %>% st_transform(3857), inherit.aes = FALSE,
            aes(fill = cut(diff_ls, breaks = bins)), lwd = 0) + 
    scale_fill_manual("Monetary Benefit (\\$)", values = mypal) +
    geom_sf(data = street_parks,  color = "black", inherit.aes = FALSE) + 
    theme(axis.line = element_line(color = NA),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          text = element_text(size=8))
    
} else {
  diff_pal <- colorBin(palette = mypal, domain = blockgroups$diff_ls, bins = bins )
  
  leaflet(blockgroups %>% st_transform(4326)) %>%
    addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
    addPolygons(fillColor = ~diff_pal(diff_ls), 
                stroke = FALSE, fillOpacity = 1) %>%
    addPolylines(data = slowstreets %>% st_transform(4326), color = "grey") %>%
    addLegend("bottomleft", pal = diff_pal, values = ~diff_ls, title = "Value of Street Opening")
}
```

More interesting than the total benefit or even its spatial distribution,
however, is the social equity of its distribution among different population
segments. If we assign the block-group level monetary benefit to each household
in the block group, we can begin to allocate the distribution of benefits
proportionally to households of different sociodemographic classifications.
Specifically, if a block group with $N$ total households has a measured consumer
surplus $\delta CS$, then the share of the total benefits going to a particular
population segment $k$ is

\begin{equation}
  S_k = N * P_k * \delta CS
  (\#eq:cs-alloc)
\end{equation}

where $P_k$ is the proportion of the block group's population in segment $k$.
There is some opportunity for confusion when some demographic variables we use
(share of households with children, household income) are defined at the
household level and others (specifically ethnicity) are defined at the person level. It is
similarly not clear whether the benefits of improved park access should be
assigned at the person level, the household level, or the number of total park
trip makers in each block group. For consistency and simplicity, we assert that
the benefit is assigned to each household, and that persons receive a
proportional share of the household benefit. For example, a block group with 30%
Black individuals will receive 30% of the benefits assigned to all the
households in the block group.

Table \@ref(tab:equity) shows the total benefit assigned to households in this
way as well as the share of all monetary benefits in the region. In some cases,
the policy of opening streets as public spaces had a pro-social benefit, as
18.7\% of benefits went to Black individuals, even though only 11.4\% of the
population of Alameda County is Black. Similarly, roughly one-quarter of total
benefits went to households making less than \$35,000 per year even though only
one-fifth of the households are in this category. On the other hand, a smaller
than expected share of benefits is allocated to Asian individuals and households
making more than $125,000 per year.

```{r equity}
tar_load(benefits)
benefits %>%
  kbl(caption = "Equity Distribution of Street Opening Benefits", booktabs = TRUE, 
      col.names = c("Group", "Benefit", "Percent* of Benefits", "Households**", 
                    "Percent of Households"),
      digits = c(0, 0, 2, 0, 2),
      align = c('l', rep('c', 4))) %>%
  kable_styling(latex_options = c("scale_down")) %>%
  column_spec(1, width = "1.8in") %>%
  column_spec(2:5, width = "1in") %>%
  footnote(symbol = c(
    "As individuals and households will belong in multiple groups, the percents do not sum to 100.",
    "Race and ethnicity are person-level attributes; households are assumed to follow the same distribution."))
```


