---
title: "OD"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tigris_use_cache = TRUE)
library(tidyverse)
library(tigris)
library(sf)
library(leaflet)
library(qwraps2)
```

In this document we examine and process the origin-destination records provided by
StreetLight Data to compile a distribution of park trip origin block groups.


The data are stored as the share of trips bound for parks (`id`) from
blockgroups (`geoid`). The `share` is the share the total `attractions` to a park
originating in that block group. There are also values for the day type and the
day share, but we use trips on all days at all times of the day.

```{r trips}
trips <- read_csv(here("data/Master_BlockGroups_Final071519.csv")) %>% 
  # need to make block group id properly formatted.
  transmute(
    id = as.character(`Zone Name`),
    geoid = paste0("0", round(`Block Group ID`)),
    county = substr(geoid, 0, 5),
    attractions = `Zone Traffic (StL Index)`,
    d_share = `Percent by Home Location`,
    volume = attractions * d_share
  ) %>%
  # only keep bgs in the area
  filter(county %in% c("06001", "06013"))

trips %>% sample_n(10)

```

Let's look at the volume of trips inbound to Lakeside Park in Oakland by Lake
Merrit from each block group by volume. Unsurprisingly, the block groups 
immediately next to the park show a great deal affinity to the park. 
But it also draws visitors from a wider group of block groups, particularly
along the east bench.

```{r lakeside-map}
block_group_shape <- tigris::block_groups(
  "CA",  county = c("001"),  class = "sf") %>%
  select(geoid = GEOID) %>%
  st_transform(4326)

inbound_trips <- block_group_shape %>% 
  left_join(
    trips %>% filter(id == "21997") %>% select(geoid, volume), 
    by = "geoid"
  ) %>%
  mutate(volume = ifelse(is.na(volume), 0, volume))

pal <- colorBin("BuPu", bins = c(0, 1, 10, 30, 50, Inf))

leaflet(inbound_trips) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(color = ~pal(volume)) %>%
  addLegend("bottomleft", pal = pal, values =~volume)
```

The `share` colum is actually the opposite of what we want to have; we are
interested not in where park visitors are coming *from*, but which parks 
people *travel to* based on their home block group.

```{r flipit}
park_flows <- trips %>%
  group_by(geoid) %>%
  mutate(
    productions = sum(volume),
    o_share = volume / productions
  )  %>%
  write_rds(here("data/park_flows.rds"))
```

The map below shows for a block group selected in black the distribution of 
observed park destinations. Interestingly enough, the nearest parks are *not*
the most commonly observed parks. But the parks that are not nearby are large
parks likely to attract people from far away.

```{r map2}
parks_map_data <- read_rds(here("data/attributed_parks.rds")) %>%
  filter(id %in% park_flows$id) %>%
  transmute(id) %>%
  left_join(park_flows %>% filter(geoid == "060014052001"), by = "id") %>%
  filter(!is.na(o_share))

leaflet(parks_map_data %>% st_transform(4326)) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(color = ~pal(volume)) %>%
  addPolygons(data = block_group_shape %>% filter(geoid == "060014052001"), 
              color = "black") %>%
  addLegend("bottomleft", pal = pal, values =~o_share)
```






